---
title: "Combinatorics Exercises"
author: "Abel Isaias Gutierrez-Cruz"
date: "25/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 5.1 
At a party with 10 guests, every guest shakes hands with each other guest. How many handshakes can be counted in total?

It does not make sense that the guests could shake hands with themselves, so replacement is not relevant. 
$\begin{pmatrix} 10 \\ 2 \end{pmatrix}$
```{r}
choose(10, 2)
```

## Exercise 5.2 
“Gobang” is a popular game in which two players set counters on a board with 381 knots. One needs to place 5 consecutive counters in a row to win the game. There are also rules on how to remove counters from the other player. Consider a match where 64 counters have already been placed on the board. How many possible combinations exist to place 64 counters on the board?

Each time that you put a counter in a knot, that position will not be taken. 
$\begin{pmatrix} 381 \\ 64 \end{pmatrix}$
```{r}
choose(381, 64)
```



## Exercise 5.3 
A shop offers a special tray of beer: “Munich’s favourites”. Customers are allowed to fill the tray, which holds 20 bottles, with any combination of Munich’s 6 most popular beers (from 6 different breweries).
a. What are the number of possible combinations to fill the tray?
The customers could choose a beer more than one time, so we should consider replacement. The order is not important. 
$n = 6$ and $m=20$
$\begin{pmatrix} n + m -1 \\ m \end{pmatrix} = \begin{pmatrix} 6 + 20 -1 \\ 20 \end{pmatrix}$
```{r}
choose(25, 20)
```

b. A customer insists of having at least one beer from each brewery in his tray. How many options does he have to fill the tray?
That customer already choose 6 positions out of 20, so we already only have 14 positions
$\begin{pmatrix} n + m -1 \\ m \end{pmatrix} = \begin{pmatrix} 6 + 14 -1 \\ 14 \end{pmatrix}$
```{r}
choose(19, 14)
```



## Exercise 5.4 
The FIFA World Cup 2018 in Russia consists of 32 teams. How many combinations for the top 3 teams exist when
a. taking into account the order of these top 3 teams and
```{r}
choose(32, 3)*factorial(3)
```

b. without taking into account the order of these top 3 teams?
```{r}
choose(32, 3)
```


## Exercise 5.5 
An online book store assigns membership codes to each member. For administrative reasons, these codes consist of four letters between “A” and “L”. A special discount period increased the total number of members from 18, 200 to 20, 500. Are there enough combinations of codes left to be assigned for the new membership codes?
```{r}
num <- length(LETTERS[LETTERS <= "L"])
num**4
```
There are enough codes


## Exercise 5.6 
In the old scoring system of ice skating (valid until 2004), each member of a jury of 9 people judged the performance of the skaters on a scale between 0 and 6. It was a decimal scale and thus scores such as 5.1 and 5.2 were possible. Calculate the number of possible score combinations from the jury.
$n=61$ and $m=9$
```{r}
61**9
```



