---
title: "Explanations in the book"
author: "Abel Isaias Gutierrez-Cruz"
date: "1/2/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## Linear Regression with Multiple Covariates
```{r}
# load data
data <- read.csv("../Data/pizza_delivery.csv")
lm(data$time~data$bill+data$operator)
```

### Linear Regression with categorical covariates
```{r}
lm(data$time~as.factor(data$branch))
summary(lm(data$time~as.factor(data$branch)))
```

### Linear regression with ANOVA
```{r}
m2 <- lm(data$time~data$branch+data$operator+data$driver)
summary(lm(data$time~data$branch+data$operator+data$driver))

anova(m2)
```

### Linear regression with categorical-continuous interaction
```{r}
int.model.1 <- data %>% lm(temperature~time*branch, .)
int.model.2 <- data %>% lm(temperature~time+branch+time:branch, .)
summary(int.model.1)
summary(int.model.2)
```

## Comparing different Models

### Backward selection

```{r}
library(MASS)
ms <- data %>% lm(time~branch+bill+operator+driver+temperature+pizzas, data= .)
stepAIC(ms, direction = 'back')
```

We obtained that removing the "operator" variable we could improve the model:
```{r}
ms <- data %>% lm(time~branch+bill+driver+temperature+pizzas, .)
stepAIC(ms, direction = 'back')
```

#### Evaluating normality assumption and homoscedasticity
```{r}
fm <- data %>% lm(time~branch+bill+driver+temperature+pizzas, .)
hist(rstandard(fm))
plot(fm, which=2)
```

In the prior plots is shown that the normality assumption does not seem to be violated

```{r}
plot(fm, which = 3)
```

The plot do not have a visible structure, there is no indication of heteroscedaticity.

